---
title: "Hit Prediction"
author: "Kat Wilson"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

I fit a two-component mixture model to the distribution of popularity averages, one that represents the normal songs, and another that represents``hit tracks". As we don't observe the grouping of songs in the data generation process (i.e. Spotify doesn't have a 'hit track' variable), then we consider this a missing data problem, and turn to the EM algorithm to identify which of the two components the data came from. A normal-normal mixture model will do well here, as the distribution of popularity appears to be bimodal. We can assume a general mixture model where the outcome is in one of two distributions:


$$
\begin{equation}
y =  Normal(\mu, \sigma^2)
\end{equation}

\begin{equation}
y =  Normal(\mu, \sigma^2)
\end{equation}
$$


In the expectation step, the expected indicator for every single observation (whether the observation is in the popular or un-popular group) is calculated. The following step takes these indicator values from the E step, and outputs new values for the parameters. In other words, the E step takes in current parameter values and outputs estimated values for the indicators. The M step takes in values for the indicators and outputs values for the parameters. I hypothesize that there bands will create less hit songs than they do regular tracks. Running the EM algorithm from these starting points, estimates result in the overall maximum log-likelihood (=360.7) with mu0 at 0.17, mu1 at .33., and alpha at 0.52. This model assigns a high probability to a song being a hit. The overlap in distributions is also troubling. By forcing the two components to have the same variance, alpha becomes 0.26, in tune with our hypothesis of hit song making being a rare event.


Figure: A histogram of the popularity averages with curves for both instances of the fitted densities using the MLE of the mixture components is below. I checked for multiple modes by running the EM algorithm starting from many different starting values.

## Vignette Info

Note the various macros within the `vignette` section of the metadata block above. These are required in order to instruct R how to build the vignette. Note that you should change the `title` field and the `\VignetteIndexEntry` to match the title of your vignette.

## Styles

The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

    output: 
      rmarkdown::html_vignette:
        css: mystyles.css

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
